---
title: "Penang fisheries"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 10, fig.height = 7)
library(ggplot2)
library(magrittr)
library(ggdist)

library(brms)

```

```{r load-data}
drake::loadd(trips)
drake::loadd(points)
drake::loadd(landings_clean)

trips <- trips %>%
  dplyr::filter(date < as.Date("2020-09-01"))
```

```{r}
full_info_trips <- dplyr::filter(trips, !is.na(trip_id_landing), !is.na(trip_id_track))

landing_trips <- dplyr::filter(trips, !is.na(trip_id_landing))
tracking_trips <- dplyr::filter(trips, !is.na(trip_id_track))

landing_fisher_tabyl <- landing_trips %>%
  janitor::tabyl(fisher)

main_contributor <- trips %>%
  dplyr::group_by(fisher) %>%
  dplyr::summarise(n_trips_landing_fisher = dplyr::n_distinct(trip_id_landing, na.rm = T), 
                   n_trips = dplyr::n(), 
                   prop = n_trips_landing_fisher/n_trips) %>% 
  dplyr::filter(n_trips_landing_fisher == max(n_trips_landing_fisher)) 

# tracking_trips %>%
#   dplyr::group_by(fisher) %>%
#   dplyr::summarise(first_trip = min(date)) %>%
#   dplyr::arrange(first_trip)
```

We report data collected between `r format(min(trips$date), "%d %B %Y")` and `r format(max(trips$date), "%d %B %Y")`. 
During this period we recorded information about `r nrow(trips)` fishing trips in Penang. 
The information about these trips was contributed by a total of `r dplyr::n_distinct(trips$fisher)` local fishers and is a combination of tracking and landings. 

During the reporting period, WorldFish installed solar-powered GPS trackers sourced from Pelagic Data Systems Inc. in `r dplyr::n_distinct(points$boat_name)` boats at four key landing sites of Teluk Bahang, Balik Pulau, Seagate and Kuala Binjai.
Most of the trackers were installed in July or August 2020, but three of them were installed in September or November 2019 which gives us a good coverage over the last year for these boats. 
Until `r format(max(trips$date), "%d %B %Y")` these units tracked a total of `r nrow(tracking_trips)` trips.  

In addition to the boat tracks, we also collected landings data informally through a WhatsApp group set up with participating fishers. 
In total, we collected landings information for `r nrow(landing_trips)` trips by `r dplyr::n_distinct(landing_trips$fisher)` fishers. 

Three of these fishers have also a tracker installed on their boats and consequently we have both tracking and landing data for `r nrow(full_info_trips)` trips.
Most of this landing data, however, comes from the most active fisher.
This fisher alone contributed landing data for `r max(landing_fisher_tabyl$n)` trips which corresponds to about `r scales::percent(main_contributor$prop[1])` of all their fishing trips over the reporting period. 

```{r}
points %>%
  dplyr::filter(time < lubridate::ymd("2020-09-01")) %>%
  dplyr::left_join(trips, by = c("trip" = "trip_id_track")) %>% 
  dplyr::group_by(trip) %>%
  dplyr::slice_sample(prop = 0.1) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(time) %>%
  dplyr::mutate(trip = forcats::fct_reorder(as.character(trip), !is.na(trip_id_landing))) %>%
  ggplot(aes(x = lng, y = lat, group = trip, colour = !is.na(trip_id_landing))) +
  geom_path(size = 0.2, alpha = 0.7) +
  guides(colour = guide_legend()) +
  scale_color_manual(values = c("grey70", "#1a4985"), 
                     labels = c("trips without landing data", "trips with landing data")) +
  coord_quickmap() +
  theme_minimal() +
  theme(legend.position = "bottom", 
        legend.title = element_blank()) +
  labs(title = "A visual sample of fishing trips in Penang", 
       subtitle = "Between September 2019-2020", 
       x = "Longitude", y = "Latitude")
```

We are interested in calculating the total catch from artisanal fisheries in Penang and the income it generates. 
We need three key components to perform these calculations. 
First, we need to estimate the expected catch weight from each trip and the income it generates. 
Second, we need to obtain an estimate of the vessel activity throughout the year. 
Third, we need an estimate of the number of fishers in Penang. 

We use hierarchical Bayesian models as the framework to estimate these figures for four main reasons. 
First, Bayesian models perform particularly well with relatively small amount of data. 
Second, a hierarchical model allow us to minimise the biases that a arise from unbalances in the contributed data; both across fishers and across time. 
Third, using models, instead of simple averages allow us to obtain estimates of the catch even for periods for which we have no comprehensive sampling. 
Lastly, and most importantly, a Bayesian framework allow us, not only to obtain an estimate of the numbers we are interested on, but also allow us to better understand the uncertainties involved in the calculation. 

```{r}
drake::loadd(landing_model)
drake::loadd(vessel_model)

models <- list(
  catch = landing_model[[1]], 
  income = landing_model[[2]], 
  vessel = vessel_model
)

predict_overall_mean <- function(model){
  model %>%
    posterior_epred(newdata = data.frame(dummy = 1), re_formula = NA) %>%
    as.data.frame() %>%
    tibble::as_tibble() %>%
    dplyr::rename(Estimate = V1)
}

est_sum <- models %>%
  purrr::map(. %>%
               predict_overall_mean() %>%
               posterior_summary()%>%
               purrr::array_tree(margin = 2))
```


First, we look at the expected catch weight from a single trip. 
There was a lot of variability in the catch weight with some trips fishing as little as `r round(min(landing_trips$weight_kg), 1)`kg and some as much as `r round(max(landing_trips$weight_kg), 0)`kg. 
Nevertheless, using ten thousand Monte Carlo simulations, we found that the average catch was somewhere between `r round(est_sum$catch$Q2.5, 1)`kg and `r round(est_sum$catch$Q97.5, 1)`kg (95% credible intervals) with a mean of `r round(est_sum$catch$Estimate, 1)`kg.
The average catch showed important differences among fishers, with some consistently landing more catch by weight than the average. 
More data from different fishers will allow us to better understand what drives these differences. 

That catch represented an average income per trip between RM`r round(est_sum$income$Q2.5)` and RM`r round(est_sum$income$Q97.5)` (mean RM`r round(est_sum$income$Estimate)`). 
Similar as the catch weight, average income levels showed a large variation among fishers and trips. 
Some trips provided only RM`r round(min(landing_trips$total_price))` as income for the fishers while some trips provided as much as RM`r round(max(landing_trips$total_price))`. 

We expect the vessel activity coefficient during the study period to be between `r round(est_sum$vessel$Q2.5, 2)` and `r round(est_sum$vessel$Q97.5, 2)` (mean `r round(est_sum$vessel$Estimate, 2)`). 
Mathematically, this coefficient can be interpreted in two equivalent ways. 
When applied to a single boat, it can be interpreted as the probability that this boat would to a fishing trip in a given day. 
When applied to a group of boats, for example the Penang artisanal fishing fleet, it can be interpreted as the proportion of boats that can be expected to be going on a trip in a given day. 
Although some differences between fishers exist, the largest variations were across time with some weeks being more . 


```{r}
set.seed(7)

predict_yearly_error <- function(model){
  pred_data <- tibble::tibble(date = seq(from = as.Date("2019-09-01"), 
                                       to = as.Date("2020-08-31"),
                                       by = 1)) %>%
  dplyr::mutate(month = lubridate::month(date), 
                week = lubridate::isoweek(date), 
                fort = as.numeric(week) %/% 2,
                wday = lubridate::wday(date, label = T))
  
  posterior_epred(
    model, 
    newdata = pred_data, nsamples = 10000, 
    re_formula = ~ (1 | wday) + (1 | week), allow_new_levels = TRUE) %>% 
    set_colnames(as.character(pred_data$date)) %>%
    as.data.frame.table() %>%
    dplyr::mutate(date = as.Date(Var2)) %>%
    dplyr::rename(Estimate = Freq, sample = Var1) %>%
    dplyr::inner_join(pred_data, by = "date") %>%
    tibble::as_tibble() %>%
    dplyr::select(-Var2)
}

label_date <- function(x){
  format(x, "%B") %>%
    stringr::str_replace("January", paste0("January\n2020"))
}

label_income <- function(x){
  x <- scales::label_dollar(prefix = "RM", scale = 1/1000000)(x)
  x[length(x):(length(x))] <- paste0(x[length(x):(length(x))], "\nmillion")
  x
}

label_weight <- function(x){
  x <- scales::label_number(scale = 1/1000, suffix = "")(x)
  x[length(x):(length(x))] <- paste0(x[length(x):(length(x))], "\ntonnes")
  x
}

posterior_draws <- models %>%
  purrr::map_dfr(predict_yearly_error, .id = "model") %>%
  tidyr::pivot_wider(id_cols = c("date", "sample", "month", "week", "wday", "fort"), 
                     names_from = "model", 
                     values_from = "Estimate") %>%
  dplyr::mutate(n_fishers = round(rnorm(dplyr::n(), 1658, 20)), 
                income = income * vessel * n_fishers,
                catch = catch * vessel * n_fishers, 
                vessel = vessel) %>% 
  dplyr::group_by(month, sample) %>%
  dplyr::summarise(date = min(date), 
                   income = sum(income), 
                   catch = sum(catch), 
                   vessel = mean(vessel))

plot_ts <- function(x, col){
  x %>% 
  ggplot(aes_string(x = "date", y = col, group = 1)) +
  stat_lineribbon(size = 0.5,
                  alpha = 1/2,
                  .width = c(.95, .66)) +
  # geom_line(aes(group = sample), size = 0.1, alpha = 0.1) +
  # stat_pointinterval(thickness = 7) +
  scale_fill_brewer(palette = "Greys") +
  expand_limits(y = 0) +
  scale_x_date(date_breaks = "month", minor_breaks = NULL, labels = label_date) +
  theme_minimal() +
    theme(axis.title = element_blank(), legend.position = "none")
}

catch_ts_plot <-
  posterior_draws %>%
  plot_ts("catch") +
    scale_y_continuous(labels = label_weight) +
    labs(title = "Monthly catch totals by artisanal fishers in Penang", 
       subtitle = "Weight")
  
income_ts_plot  <- 
  posterior_draws %>%
  plot_ts("income") +
    scale_y_continuous(labels = label_income) +
  labs(subtitle = "Income",
       caption = "Shaded areas indicate the 66 and 95% credible intervals")

vessel_ts_plot <- 
  posterior_draws %>%
  plot_ts("vessel")  +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(subtitle = "Vessel activity")
# 
# title_plot <- ggplot() +
#   labs(title = "Time series") +
#   theme_minimal()
cowplot::plot_grid(#title_plot, 
                   catch_ts_plot, 
                   income_ts_plot, 
                   rel_heights = c(1.05, 1),
                   # vessel_ts_plot,
                   ncol = 1, align = "v")
```

```{r}
annual_totals <- posterior_draws %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(dplyr::across(tidyselect:::where(is.numeric), sum))

ann_sum <- annual_totals %>%
  dplyr::select_if(is.numeric) %>%
  posterior_summary() %>%
  set_names(NULL) %>%
  purrr::array_tree()

```

Using all these pieces of information we estimate that between the period `r format(min(trips$date), "%d %B %Y")` and `r format(max(trips$date), "%d %B %Y")` **artisanal fishers were very likely (95% probability) to catch between `r round(ann_sum$catch$Q2.5/1000000, 1)` and `r round(ann_sum$catch$Q97.5/1000000, 1)` thousand tonnes in Penang waters, which provided income between RM`r round(ann_sum$income$Q2.5/1000000, 0)` and RM`r round(ann_sum$income$Q97.5/1000000, 0)` million to the local communities**. 

```{r, fig.height=5}
annual_catch_dist_plot <- annual_totals %>%
  dplyr::slice_sample(n = 1000, replace = F) %>%
  ggplot(aes(x = catch)) +
  stat_dotsinterval(.width = c(.95, .66), fill = "grey90", slab_colour = "grey90") +
  geom_text(aes(x = mean(catch), 
                label = stat(paste0("mean: ", round(mean(x)/1000000, 1), " thousand tonnes")), 
                y = 0, group = 1), stat = "unique", vjust = -1) +
  scale_x_continuous(labels = scales::label_number(scale = 1/1000000, suffix = " thousand\ntonnes", accuracy = 1)) +
  theme_minimal() +
  theme(axis.title = element_blank(), 
        axis.text.y = element_blank(), 
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(title = "Estimated annual catch by artisanal fishers in Penang", 
       subtitle = paste("Weight"))

annual_income_dist_plot <- annual_totals %>%
  dplyr::slice_sample(n = 1000, replace = F) %>%
  ggplot(aes(x = income)) +
  stat_dotsinterval( .width = c(.95, .66), fill = "grey90", slab_colour = "grey90") +
  geom_text(aes(x = mean(income), 
                label = stat(paste0("mean: RM", round(mean(x)/1000000), " million")), 
                y = 0, group = 1), stat = "unique", vjust = -1) +
  scale_x_continuous(labels = scales::label_dollar(prefix = "RM", scale = 1/1000000, suffix = "\nmillion")) +
  theme_minimal() +
  theme(axis.title = element_blank(), 
        axis.text.y = element_blank(), 
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),) +
  labs(subtitle = "Income", 
       caption = "* Each dot corresponds to a simulation in our model. Horizontal lines indicate the 66 and 95% credible intervals")

cowplot::plot_grid(annual_catch_dist_plot, 
                   annual_income_dist_plot, 
                   ncol = 1,
                   align = "v")
```


```{r eval = F}
drake::loadd(landing_model)
drake::loadd(vessel_model)

landing_model[[1]] %>%
  predict_overall_mean() %>%
  posterior_summary()




vessel_model %>%
  posterior_samples(pars = "^Intercept") %>%
  as.data.frame() %>%
  tibble::as.tibble() %>%
  dplyr::mutate(Estimate = plogis(Intercept)) %>%
  ggplot(aes(x = Estimate, y = 0)) +
  stat_halfeye() +
  theme_minimal()

landing_model[[2]] %>%
  posterior_samples(pars = c("^Intercept", "sigma")) %>%
  as.data.frame() %>%
  tibble::as.tibble() %>%
  dplyr::mutate(Estimate = exp(Intercept + (sigma)/2), m = mean(Estimate)) %>%
  ggplot(aes(x = Estimate, y = 0)) +
  stat_halfeye() +
  theme_minimal()

landing_model[[2]] %>%
  posterior_predict(newdata = data.frame(dummy = 1), re_formula = NA) %>%
  as.data.frame() %>%
  tibble::as.tibble() %>%
  dplyr::mutate(Estimate = V1, 
                m = mean(Estimate)) %>%
  ggplot(aes(x = Estimate, y = 0)) +
  stat_halfeye() +
  theme_minimal()

predict_yearly <- function(model){
  pred_data <- tibble::tibble(date = seq(from = as.Date("2019-09-01"), 
                                       to = as.Date("2020-08-31"),
                                       by = 1)) %>%
  dplyr::mutate(month = lubridate::month(date), 
                week = lubridate::isoweek(date), 
                wday = lubridate::wday(date, label = T))
  
  predict(model, pred_data, re_formula = ~ (1 | wday) + (1 | week), allow_new_levels = TRUE) %>%
    as.data.frame() %>%
    dplyr::bind_cols(pred_data) 
}

plot_time_prediction <- function(x) {
  x %>%
    ggplot(aes(x = date, y = Estimate)) +
    geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5), fill = "grey70") +
    geom_line() +
    theme_minimal()
}

predict_yearly(vessel_model) %>%
  stat_dotsinterval()
```


```{r, eval = FALSE}
library(brms)
landing_model_data <- landing_trips %>%
  dplyr::ungroup() %>%
  dplyr::mutate(month = lubridate::month(date), 
                week = lubridate::isoweek(date), 
                wday = lubridate::wday(date, label = T), 
                num_date = as.numeric(date), 
                num_date = num_date - min(num_date)) %>%
  dplyr::group_by(fisher, date) %>%
  dplyr::slice(1)

m0 <- brm(weight_kg ~ 1 + (1 | fisher) + (1 | wday) + (1 | month), 
          data = landing_model_data, 
          family = lognormal, 
          iter = 4000, 
          save_pars = save_pars(all = TRUE),
          cores = 4)

m2 <- brm(weight_kg ~ 1 + arma(time = num_date, gr = fisher, cov = TRUE) + (1 | fisher) + (1 | wday) + (1 | week), 
          data = landing_model_data, 
          family = lognormal, 
          control = list(adapt_delta = 0.9),
          save_pars = save_pars(all = TRUE),
          iter = 4000, 
          cores = 4)

pred_data <- tibble::tibble(date = seq(from = as.Date("2019-09-01"), 
                                       to = as.Date("2020-08-31"),
                                       by = 1)) %>%
  dplyr::mutate(month = lubridate::month(date), 
                week = lubridate::isoweek(date), 
                wday = lubridate::wday(date, label = T))

plot_time_prediction <- function(x) {
  x %>%
    ggplot(aes(x = date, y = Estimate)) +
    geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5), fill = "grey70") +
    geom_line() +
    theme_minimal()
}

samples_to_df <- . %>%
  set_colnames(as.character(pred_data$date)) %>%
  as.data.frame.table() %>%
  dplyr::mutate(date = as.Date(Var2)) %>%
  dplyr::rename(Estimate = Freq, sample = Var1) %>%
  dplyr::inner_join(pred_data, by = "date")

m0_pred_summary <- predict(m0, pred_data, re_formula = ~ (1 | wday) + (1 | month), allow_new_levels = TRUE) %>%
  as.data.frame() %>%
  dplyr::bind_cols(pred_data) 

m0_pred_samples <- predict(m0, pred_data, re_formula = ~ (1 | wday) + (1 | month),
                           allow_new_levels = TRUE, summary = F) %>%
  samples_to_df()

m1_pred_summary <- predict(m1, pred_data, re_formula = ~ (1 | wday) + (1 | week), allow_new_levels = TRUE) %>%
  as.data.frame() %>%
  dplyr::bind_cols(pred_data) 

m1_pred_samples <- predict(m1, pred_data, re_formula = ~ (1 | wday) + (1 | week), allow_new_levels = TRUE, 
                           summary = FALSE)  %>%
  samples_to_df()


m0_pred_samples %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(yearly_total = sum(Estimate)) %>%
  dplyr::mutate(group = rep(c("A", "B"), each = 4000)) %>%
  ggplot(aes(x = yearly_total, y = 1)) +
  stat_dotsinterval()
```


## Landings




```{r}
```

```{r, message=FALSE}
trip_summary <- landings_clean %>%
  dplyr::group_by(trip_id) %>%
  dplyr::summarise(
    date = dplyr::first(date), 
    fisher = dplyr::first(fisher), 
    total_price = sum(total_price), 
    weight_kg = sum(weight_kg),
    n_species = dplyr::n_distinct(common_malay))

fisher_summary <- landings_clean %>%
  dplyr::group_by(fisher) %>%
  dplyr::summarise(
    imei = dplyr::first(imei),
    first_trip = min(date), 
    last_trip = max(date),
    n_trips = dplyr::n_distinct(trip_id),
    total_price = sum(total_price), 
    weight_kg = sum(weight_kg),
    n_species = dplyr::n_distinct(common_malay)) %>%
  dplyr::arrange(-n_trips)
```

Between `r format(min(trip_summary$date), "%d %B %Y")` and `r format(max(trip_summary$date), "%d %B %Y")` we recorded landings for `r nrow(trip_summary)` trips by `r dplyr::n_distinct(landings_clean$fisher)` fishers. 



During this time, GPS trackers were installed in `r length(na.omit(fisher_summary$imei))` boats. 

```{r}
trip_summary %>%
  ggplot(aes(x = date)) +
  geom_histogram(binwidth = 14, boundary = as.Date("2020-01-01")) +
  theme_minimal()
```

```{r}
fisher_summary %>%
  dplyr::mutate(fisher = forcats::fct_reorder(fisher, first_trip, .desc = TRUE)) %>%
  ggplot(aes(y = fisher)) +
  geom_segment(aes(x = first_trip, 
                   xend = last_trip, yend = fisher, colour = is.na(imei))) +
  geom_point(data = trip_summary, 
             aes(x = date), size = 1, shape = 3) +
  theme_minimal()
```

```{r}
species_summary <- landings_clean %>%
  dplyr::filter(!is.na(price_kg)) %>%
  dplyr::mutate(species = forcats::fct_lump_n(common_malay, w = price_kg*weight_kg, n = 20)) %>%
  dplyr::group_by(species) %>%
  dplyr::summarise(
    frequency = dplyr::n_distinct(trip_id), 
    weight = sum(weight_kg, na.rm = TRUE), 
    price_kg = mean(price_kg, na.rm = TRUE)) %>%
  dplyr::mutate(species = forcats::fct_reorder(species, price_kg*weight), 
                species = forcats::fct_relevel(species, "Other"))

species_summary %>%
  ggplot(aes(x = species, y = price_kg*weight)) +
  geom_col() +
  coord_flip()

species_summary %>%
  ggplot(aes(x = species, y = price_kg)) +
  geom_col() +
  coord_flip()

species_summary %>%
  ggplot(aes(x = species, y = weight)) +
  geom_col() +
  coord_flip()
```

```{r}
trip_summary %>%
  ggplot(aes(x = date, y = weight_kg)) +
  geom_point(aes(colour = fisher))

trip_summary %>%
  ggplot(aes(x = date, y = total_price)) +
  geom_point(aes(colour = fisher))
```


```{r}
predict_fisher_differences <- function(model){
  model %>%
    posterior_epred(newdata = data.frame(fisher = unique(model$data$fisher)), re_formula = ~ (1 | fisher)) %>%
    set_colnames(unique(model$data$fisher)) %>%
    as.data.frame.table() %>%
    dplyr::rename(fisher = Var2) %>%
    dplyr::rename(Estimate = Freq, sample = Var1)
}

models %>%
  purrr::map_dfr(.f = predict_fisher_differences, .id = "model") %>%
  dplyr::mutate(
    # Randomise names to protect identity
    fisher = factor(
      x = fisher, 
      labels =  randomNames::randomNames(
        n = dplyr::n_distinct(fisher), 
        ethnicity = 6, 
        which.names = "first", 
        sample.with.replacement = FALSE)),
    # Order fishers by catch weight
    fisher = forcats::fct_reorder2(
      .f = fisher, 
      .x = Estimate, 
      .y = as.numeric(model == "catch"), 
      .fun = weighted.mean, 
      .desc = TRUE), 
    # Reverse order so that fishers without catch show at the bottom of the plot
    fisher = forcats::fct_rev(f = fisher),
    model = forcats::fct_relevel(model, "catch")) %>%
  ggplot(aes(x = Estimate, y = fisher)) +
  stat_pointinterval(point_size = 1.5) +
  facet_wrap("model", scales = "free_x") +
  theme_minimal() +
  labs(
    title = "There were important differences among fishers",
    caption = "* Fisher names have been changed to protect their identity")

predict_wday_differences <- function(model){
  model %>%
    posterior_epred(newdata = data.frame(wday = unique(model$data$wday)), re_formula = ~ (1 | wday)) %>%
    set_colnames(unique(model$data$wday)) %>%
    as.data.frame.table() %>%
    dplyr::rename(wday = Var2) %>%
    dplyr::rename(Estimate = Freq, sample = Var1)
}

models %>%
  purrr::map_dfr(.f = predict_wday_differences, .id = "model")%>%
  ggplot(aes(x = Estimate, y = wday)) +
  stat_pointinterval() +
  facet_wrap("model", scales = "free_x")

predict_week_differences <- function(model){
  model %>%
    posterior_epred(newdata = data.frame(week = unique(model$data$week)), re_formula = ~ (1 | week)) %>%
    set_colnames(unique(model$data$week)) %>%
    as.data.frame.table() %>%
    dplyr::rename(week = Var2) %>%
    dplyr::rename(Estimate = Freq, sample = Var1)
}

models %>%
  purrr::map_dfr(.f = predict_week_differences, .id = "model") %>%
  dplyr::mutate(week = forcats::fct_reorder(week, as.numeric(as.character(week)))) %>%
  ggplot(aes(x = Estimate, y = week)) +
  stat_pointinterval() +
  facet_wrap("model", scales = "free_x")
```

